---
title: "Qf LBL"
author: "Kelsey Walker"
date: "12/21/2021"
output: html_document
---

---
copied from "LBL_Model" in Fall-2021 folder
"10/28/2021"
---
# Packages
```{r include = FALSE}
library("dplyr")
library("tidyverse")
library("readxl")
library("ggplot2")
library("lubridate")
library("gridExtra")
```

# Set Constant Parameters

#### Assumed or given values
```{r}
R <- 0.5 # even leakage btw floor and ceiling
X <- 0 # leakage distriubtion parameter

C <- 0.102 #sheilding coefficient class 5 

gamma_h <- 0.35 #house terrain class 5
alpha_h <- 0.47 #house terrain class 5
gamma_a <- 0.15 #airport terrain class 2
alpha_a <- 1.00 #airport terrain class 2

h_a <- 20*0.3048 #height of (KFNL) airport wind tower [m]
g <- 9.81 #gravity (m/s2)
p_air <- 1.225 #outdoor density of air (kg/m3). 1.225 is for sea level at 15deg C. In FoCo, should be higher (due to lower humidity and altitude)  ?????????????????
```


# Read in data

#### House Assessment Data
```{r include = FALSE}
assessment <- read_excel("~kelseywalker/Desktop/EPIC HOMES/house_characteristics_assessment/h_charact.xlsx", sheet = "LBL Model") 
  
# Define ela [m^2] and h_h [m]
## Assume each floor = 9ft, so height = 9*floor* 0.3048 m
assessment <- mutate(assessment, height = floors * 9*0.3048) %>%
  mutate(ela = ela * 1/1550) %>% # 1m^2 = 1550 in^2
  drop_na(ela, floors) %>% #filter to only homes with both ela and floor data
  rename(id_home = home)
```

#### Wind Data (Hourly)
```{r include = FALSE}
# Hourly Wind Data
wind_data <- read_excel("~kelseywalker/Desktop/EPIC HOMES/infiltration/wind_speed/FCLA_wind.xlsx", sheet = "wind_r", skip = 3)

#Remove wind_dir from dataset
wind_data <- select(wind_data, !c("Wind Dir (Deg)")) %>%

#Round time to hour and define wind speed
  mutate(
    datetime = ymd_hms(wind_data$datetime),
    datetime = round_date(datetime, 'hour') ) %>%
    rename(wind_va = 'Wind Speed (MPH)') %>%
    mutate(wind_va = wind_va * 0.44704) %>% #wind speed, [m/s]
    drop_na(wind_va) 
    
# Remove duplicate datetime (keep last)
 wind_data <- wind_data[!duplicated(wind_data$datetime),fromLast = TRUE] 
 
#ready for append process
 wind_data <- mutate(wind_data, id_home = '001')

```

```{r, include = FALSE}
#append process
wind_long <- (rbind(wind_data, mutate(wind_data, id_home = '002') )) %>%
  rbind(wind_data, mutate(wind_data, id_home = '003') )%>%
  rbind(wind_data, mutate(wind_data, id_home = '005') )%>%
  rbind(wind_data, mutate(wind_data, id_home = '006') )%>%
  rbind(wind_data, mutate(wind_data, id_home = '007') )%>%
  rbind(wind_data, mutate(wind_data, id_home = '010') )%>%
  rbind(wind_data, mutate(wind_data, id_home = '011') )%>%
  rbind(wind_data, mutate(wind_data, id_home = '012') )%>%
  rbind(wind_data, mutate(wind_data, id_home = '013') )%>%
  rbind(wind_data, mutate(wind_data, id_home = '014') )%>%
  rbind(wind_data, mutate(wind_data, id_home = '015') )%>%
  relocate(id_home, .before = datetime)
  
```

###### Combine Assessment data and Wind Data
```{r}
wind_all <- merge(wind_long, assessment, by = "id_home") %>%
  select(!c(round, floors)) %>%
  mutate(wind_vh = wind_va *( (alpha_h*(height/10)^(gamma_h)) / (alpha_a *(h_a / 10)^gamma_a) ) ) %>%
  relocate(ela, .after = wind_vh)
  
```


#### Temperature Data (Hourly)
```{r include = FALSE}
# Temperature Data 

temp_all <- readRDS("~kelseywalker/Desktop/EPIC HOMES/Fall-2021/omni_1hour.rds") %>%
  filter(location == 'living' | location == 'outdoor') %>% 
  select(id_home, datetime, location, temp) %>%
  mutate(temp = temp + 273.15 ) %>% # Convert to Kelvin
  filter(temp < 323.15) #filter: Keep only T< 50 C (122F)

## Reshape temp_all from long to wide
temp_wide <- pivot_wider(data = temp_all,
                         names_from = location,
                         values_from = temp)
temp_wide <- drop_na(temp_wide)
                      
```

#### Combine temp and windspeed using datetime and id_house as unique keys 
```{r include = FALSE}
inputs <- inner_join(wind_all, temp_wide, by = c('id_home', 'datetime'))%>%
  mutate(delta_t = living - outdoor) 

```


# LBL Model

#### Stack Effect
```{r}
outputs <- mutate(inputs, Qs = ela / 3 * sqrt(g*height * abs(delta_t)/living) * (1+R/2) * 1)  # [m^3/s} #last term = 1 = (1-X^2/(2-R)^2)^(3/2)
```

#### Wind Effect (Modified LBL Model)
```{r}
outputs <- mutate(outputs, Qw = ela * wind_vh * (1-R)^(0.8) * C) # [m^3/s]
```

#### Combined Stack and Wind
```{r}
outputs <- mutate(outputs, Qf_m3s = sqrt( Qs^2 + Qw^2)) %>% # [m^3/s] 
  mutate(Qf_cfm = Qf_m3s * 2118.88) #[ft^3/min]
```

#### WRite Output to csv 
```{r}
write_csv(outputs, "lbl_model.csv")
```

# Graphs

## results - Re-arrange output data...DONT NEED TO DO!
```{r}
###CHECK BACK. SHOULD BE ABLE TO USE pivot_wider with DATETIME instead of n. Shouldn't have duplicate issue?????????????????????????????

#outputs <- mutate(outputs, n = 1:nrow(outputs))
#results2 <- select(outputs, id_home, n, Qf_cfs) %>%
  #arrange(n) %>%
  #pivot_wider(names_from = id_home, values_from = Qf_cfs)

#results <- spread(outputs, key = id_home, value = Qf_cfs) %>%
  #select(!c(wind_va, height, wind_vh, ela, living, outdoor, delta_t, Qs, Qw, Qf_m3s, n ))

```

## Total Data Graphs and Stats
```{r}

```

## Individual House Graphs

#### Time Series Plots
```{r}

p1 <- ggplot(data = filter(outputs, id_home == "001"), aes(datetime, Qf_cfm)) + geom_line() +
 ggtitle("House 001") 

p2 <- ggplot(data = filter(outputs, id_home == "002"), aes(datetime, Qf_cfm)) + geom_line() +
 ggtitle("House 002") 

p3 <- ggplot(data = filter(outputs, id_home == "003"), aes(datetime, Qf_cfm)) + geom_line() +
 ggtitle("House 003") 

p5 <- ggplot(data = filter(outputs, id_home == "005"), aes(datetime, Qf_cfm)) + geom_line() +
 ggtitle("House 005") 

p6 <- ggplot(data = filter(outputs, id_home == "006"), aes(datetime, Qf_cfm)) + geom_line() +
 ggtitle("House 006")  

p7 <- ggplot(data = filter(outputs, id_home == "007"), aes(datetime, Qf_cfm)) + geom_line() +
 ggtitle("House 007") 

p10 <- ggplot(data = filter(outputs, id_home == "010"), aes(datetime, Qf_cfm)) + geom_line() +
 ggtitle("House 010") 

p11 <- ggplot(data = filter(outputs, id_home == "011"), aes(datetime, Qf_cfm)) + geom_line() +
 ggtitle("House 011") 

p12 <- ggplot(data = filter(outputs, id_home == "012"), aes(datetime, Qf_cfm)) + geom_line() +
 ggtitle("House 012") 

p14 <- ggplot(data = filter(outputs, id_home == "014"), aes(datetime, Qf_cfm)) + geom_line() +
 ggtitle("House 014") 

p15 <- ggplot(data = filter(outputs, id_home == "015"), aes(datetime, Qf_cfm)) + geom_line() +
 ggtitle("House 015") 

p_houses <- grid.arrange(p1, p2, p3, p5, p6, p7, p10, p11, p12, p14, p15, nrow = 3)
```


```{r, include = FALSE}
###### Save above Time Series Plots (pdf)
ggsave(filename = "lbl_individual_houses_line.pdf", plot = p_houses)
```

#### Contributing Qs and Qw 
```{r}
#1 3 6 12
 q1 <- ggplot(data = filter(outputs, id_home == "001"), aes(datetime, Qs, colour = "Qs")) + geom_line() +
  geom_line(data = filter(outputs, id_home == "001"), aes(datetime, Qw, colour = "Qw")) +
  geom_line(data = filter(outputs, id_home == "001"), aes(datetime, Qf_m3s, colour = "Qf"), alpha = 0.3) +
 ggtitle("House 001: sqrt(Qs^2 + Qw^2) = Qf")
  
 q3 <- ggplot(data = filter(outputs, id_home == "003"), aes(datetime, Qs, colour = "Qs")) + geom_line() +
  geom_line(data = filter(outputs, id_home == "003"), aes(datetime, Qw, colour = "Qw")) +
  geom_line(data = filter(outputs, id_home == "003"), aes(datetime, Qf_m3s, colour = "Qf"), alpha = 0.3) +
 ggtitle("House 003: sqrt(Qs^2 + Qw^2) = Qf")
 
 q7 <- ggplot(data = filter(outputs, id_home == "007"), aes(datetime, Qs, colour = "Qs")) + geom_line() +
  geom_line(data = filter(outputs, id_home == "007"), aes(datetime, Qw, colour = "Qw")) +
  geom_line(data = filter(outputs, id_home == "007"), aes(datetime, Qf_m3s, colour = "Qf"), alpha = 0.3) +
 ggtitle("House 007: sqrt(Qs^2 + Qw^2) = Qf")
 
 q12 <- ggplot(data = filter(outputs, id_home == "012"), aes(datetime, Qs, colour = "Qs")) + geom_line() +
  geom_line(data = filter(outputs, id_home == "012"), aes(datetime, Qw, colour = "Qw")) +
  geom_line(data = filter(outputs, id_home == "012"), aes(datetime, Qf_m3s, colour = "Qf"), alpha = 0.3) +
 ggtitle("House 012: sqrt(Qs^2 + Qw^2) = Qf")

q_houses <- grid.arrange(q1, q3, q7, q12, nrow = 2)
```

```{r, include = FALSE}
ggsave(filename = "lbl_individual_houses_Qs_Qw.pdf", plot = q_houses)
```

###### Box Plots

```{r}
#Sample Size to include on x-axis
sample_size = outputs %>%
  group_by(id_home) %>%
  summarize(num = n())


box_plot <- merge(outputs, sample_size, by = "id_home") %>%
  mutate(myaxis = paste0(id_home, "\n", "n=", num))

b_houses <- ggplot(data = box_plot, aes(x = myaxis, y = Qf_cfm)) +
geom_violin(width = 1) +
  geom_boxplot(width = 0.1, color = "grey", alpha = 0.2) +
  ggtitle("House violin plots") +
  labs(x = "House ID") +
  labs(y = "Qf (cfs)")
```


```{r, include = FALSE}
ggsave(filename = "lbl_box_plot.pdf", plot = b_houses )
```
